---
title: "Toboggan Creek Coho Smolt Program 2024"
author: "Annie Kim and Kristen Peck, DFO North Coast Stock Assessment Branch"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  word_document:
    reference_docx: "Template_doc.docx"
    toc: yes
    toc_depth: 3
---
\newpage


# **ABSTRACT**

Kim, A. and Peck, K. 2026. Toboggan Creek Coho Smolt Program 2024. Can. Manuscr. Rep. Fish. Aquat. Sci. XXXX: XX + XX p.

The Toboggan Creek Coho Smolt Program is part of the longest running coho indicator program in the North Coast Area of British Columbia. Following an unfortunate whole-cohort failure of marked hatchery smolt releases at Toboggan Creek Hatchery, a unique opportunity arose to implement a wild smolt marking program. This program ran in the spring of 2024 to estimate the abundance, migration timing, and characteristics of wild smolts emigrating from Toboggan Creek, as well as mark a large component of the run with coded-wire tags (CWTs). Wild coho smolts were captured using Rotary Screw Traps (RSTs) at two sites between the first of May and late June. A total of 18,615 wild coho smolts were captured, with a subset biosampled for size, weight, age, and condition. Smolts were adipose clipped and coded-wire tagged, with tag retention evaluated through overnight holding procedures. Peak outmigration occurred from late May to mid-June. Age analysis indicated that the population was dominated by age two smolts, followed by age three and age four smolts (Gilbert-Rich ages). Tag retention rates were high with a total of 15,420 wild smolts estimated to be released with CWTs. The population was estimated at 35,717 (1,317 SE) individuals migrating from Toboggan Creek. Overall, the 2024 program successfully met its objectives and provided valuable information on wild coho smolts in Toboggan Creek. Continued monitoring will be important for evaluating assumptions about survival between wild and hatchery coho and for improving understanding of outmigration patterns and marine exploitation of coho in northern British Columbia.  

# **RÉSUMÉ**

Kim, A. and Peck, K. 2026. Toboggan Creek Coho Smolt Program 2024. Can. Manuscr. Rep. Fish. Aquat. Sci. XXXX: XX + XX p.

\newpage

# **1. INTRODUCTION**

Toboggan Creek, situated near Smithers, British Columbia, is a glacial tributary to the Bulkley River that flows into the Skeena River. This system sustains a coho salmon (*Oncorhynchus kisutch*, “coho”) population that has been augmented by a hatchery since 1987. Since its inception, the Toboggan Creek Salmon and Steelhead Enhancement Society (TCSSES) has overseen the hatchery operations and conducted most of the stock assessment activities tied to the program. The Toboggan Creek Coho Indicator Program has offered the longest standing indicator program for North Coast British Columbia and has been crucial for assessing stock status (PSC NBTC, 2022). This program includes both a smolt and adult program, providing vital trend information about smolt production, harvest rates, marine survival, and annual escapement.

The Toboggan Creek Coho Smolt Program, which began in 1995 by DFO and was delivered by SKR Consultants, was designed to collect data on the number and age distribution of coho smolts migrating from the creek. A fyke trap was initially used until  the transition to a  Rotary Screw Trap (RST) in 1998 to reduce stress and mortality. In 1999, a wolf type weir (Wolf, 1951) was constructed upstream of the Toboggan Creek hatchery to allow greater catch of coho and upstream passage of steelhead trout (*O. mykiss*). In 2000,  an inclined plane trap (IPT) was installed downstream of the hatchery and used in conjunction to the wolf type weir upstream (SKR Consultants, 2001).

The project ran sporadically in the following years and became consistent by DFO in 2020, operated by the TCSSES. Marked hatchery smolts volitionally out-migrated from the hatchery ponds and were used as the “marked” component of the mark-recapture program estimating the smolt population size. An advantage of using Coded Wire Tags (CWT) was the ability to draw information on survival and/or abundance of a specific stock from highly mixed-stock fisheries (PSC NBTC, 2002). This method was particularly strong in predicting marine survival for tagged smolts. Since the Toboggan Creek Coho Smolt Program serves as an indicator for wild coho populations, it has operated under the assumption that hatchery and wild-origin smolts follow similar trends of migration and survival. However, whether survival and exploitation rates of hatchery and natural origin coho are the same is an assumption that has been recently called under question  (Unpublished data, DFO Smithers, 2024). Data from 2020 to 2023 suggest that survival rates may differ between wild and hatchery-origin coho, with environmental factors, management practices, and biological characteristics potentially influencing survival outcomes  (Unpublished data, DFO Smithers, 2024).

The absence of a CWT hatchery smolt release in the spring of 2024, due to a disease outbreak at the hatchery, created challenges for determining exploitation rates for that year class. However, this setback presented a valuable opportunity to focus efforts on a wild-marking initiative. Therefore, DFO Stock Assessment ran a wild smolt trapping program in spring 2024 to bridge the gap of this missing year.

## **1.1 Objectives**

**1.	Estimate the wild smolt population size out-migrating from Toboggan Creek using mark-recapture.**

**2.	Characterize size and age composition of the wild smolt outmigration from Toboggan Creek.**

**3.	Apply a large number of marks (adipose clip and coded-wire tag) to assess marine exploitation.**

\newpage

# **2. METHODS**

The Toboggan Smolt Coho Smolt Program required intensive labour, especially during peak migration, high water, and debris events. RSTs are large, floating cylindrical traps that, when positioned in the thalweg of a stream, funnel and capture juvenile fish as they migrate downstream. For the purpose of this project, RSTs were set to fish overnight to intercept periods of peak smolt activity. Continuous rotation was crucial for trapping success, and when debris obstructed rotation, entire nights of trapping were lost as the lack of rotation allowed fish to swim out of the trap box. Regular trap checks included inspections of the anchor system, connections, and pulleys to ensure they were in good condition and securely fastened. The Toboggan creek RSTs were prone to debris accumulation in the live boxes from leaves, logs, and branches, with a potential for high mortality rates or escape of juvenile fish if left unchecked. Therefore, early checks and working in split shifts was required. The programs ran five days per week from Sunday evening to Friday morning during the start and end of the migration period when fewer smolts were intercepted. During the peak migration, shifts ran seven days per week. The shifts were split into two per day: a morning shift beginning at 5 am and ending once all the fish from the two traps were processed, and an evening shift at 8 pm for setting the traps. Traps were set overnight only because this was the time when the majority of the smolts out-migrated (data on file, DFO Smithers). On most days, three crew members were required for daily operations.

During peak migration and high water events, the number of visits were dependent on the conditions of weather, forecast, and timing. Smolt migration timing in 2024 was similar to previous years where fish migration intensity increased after the second week of May and peaked from the end of May to mid-June.

Two sites (**Figure 1**)  were selected for the operation of this program, both accessed through private land:


•	Red Barn: Located 2.5 km downstream of the hatchery. A 4-ft diameter RST was installed at this site.

•	Adult Fence: The adult fence site was located 4.5 km downstream of the hatchery. Outside of the adult enumeration season in late summer and fall, a wooden sill with no adult fish migration barrier exists at this location. A 6-ft diameter RST was installed directly upstream of this structure and downstream of a private bridge. This was the same location as used for the previous years of smolt trapping but a larger diameter trap was installed in 2024 to maximize capture rates.

-add map-

*Figure 1. Map of the Toboggan Creek Coho Smolt Study Area. The two RST locations, the Red Barn and the Adult Fence, are 2.5 km and 4.5 km downstream of the Toboggan Creek Hatchery, respectively. Map created on QGIS by R. Hugdahl, DFO, 2026. Base map: Coordinate reference system (CRS) is WGS 84 (EPSG:4326). Shapefiles used: Province of British Columbia, [Freshwater Atlas Rivers], DataBC. Licensed under the Open Government Licence – British Columbia; and Province of British Columbia, [BC Major Cities Points 1:2,000,000 (Digital Baseline Mapping)]. DataBC, Licensed under the Open Government Licence – British Columbia.*

## **2.1 ENVIRONMENTAL CONDITIONS**

To investigate the relationship between environmental variables and coho smolt outmigration, we recorded discharge and temperature data at each trap check. Discharge data was obtained from the Water Survey of Canada’s real-time hydrometric station on Simpson Creek at the Mouth (station 08EE012; Environment Canada 2025), which served as a proxy for Toboggan Creek in the absence of a direct gauge. This substitution was deemed appropriate through DFO staff observation of flow patterns during field visits. Water temperatures and levels were measured and recorded by DFO technicians at the start of each site visit for both the Red Barn and Fence sites.

## **2.2 BIOSAMPLING AND CWT TAGGING PROCEDURES**

Detailed methods are found in Chicatun and Peck (2024), but relevant details are included below. While operating the traps, the goal was to minimize fish mortality and protect infrastructure. Both sites carried out adipose clipping and CWT marking with standard 1.1 mm tags with several codes. Additionally, a weekly-stratified clip was applied to smolts at the Red Barn (upstream) trap to estimate weekly outmigration, alternating between an upper and lower caudal clip each week. Recaptured fish at the downstream trap were inspected for a caudal clip. Biosampling of up to 25 smolts from each trap per day, including measurements of fork length, weight, and scale samples was conducted at both sites.

The first 25 coho smolts were measured for fork length to the nearest millimeter, and weights to the nearest 0.1 g. Scales were sampled starting at a fork length of 60 mm, with a target of 5-10 fish sampled for each 5 mm fork length interval. Any exceptionally large or small fish (down to 60 mm) was sampled for scales. Scales were sent to DFO’s Sclerochronology Lab at the end of the season for age determination in following the juvenile age sampling procedure (MacLellan 2004). All wild smolts >=70 mm were adipose clipped and tagged. Fish <70 mm were transferred to a separate bucket and released. This size cut off for coded-wire tagging was based on previous sampling data in Toboggan where the majority of smolts were >=70 mm. Similar coho smolt programs have also tagged smolts >=70 mm or >=80 mm (Leon et al., 2009 and Van et al., 2010) and considered coho smaller than that to be pre-smolts and fry (Noble et al., 2020).  At the Red Barn site, a caudal clip was applied according to the weekly stratification starting on Mondays (top or bottom). Fish were sorted based on size and tagged using the appropriate head mold size: small fish (70-85 mm): 65/lb head mold, big fish (86-115 mm): 45/lb, and bigger and biggest fish (116-190 mm or >190 mm): 15/lb head mold. Whenever there was a new tagger or tagging machine used, particularly at the beginning of the season, a small number of smolts were tagged and sacrificed to check for correct tag placement.

Tricaine methanesulfonate (TMS or MS222), diluted to 60mg/L, was used to sedate smolts prior to handling. A 5-gallon bucket and medium-sized dip nets were used to transport the fish from the RST holding box to the biosampling area. Battery-operated bubblers with diffusion stones were placed in the buckets to provide aeration for the waiting fish. A V-detector (Northwest Marine Technologies) was used to check for CWT retention and tagged smolts were tallied by tag code.

After the first 25 coho, remaining coho were visually sorted into the previously mentioned size categories for tagging. Every fish tagged was immediately checked for retention using the V-detector and placed in a recovery bucket. Once all the fish were tagged and recovered from the anesthetic, they were released in a calm eddy downstream. Any non-target fish were identified and, if time permitted, measured for length and weight.

Unexpectedly, a small number of adipose clipped hatchery coho from previous years were caught and identified by their healed-over adipose-clip, checked for CWT retention, measured and recorded, and sacrificed to check the tag code. Hatchery fish were frozen in Ziplock bags, and later dissected for CWT retrieval. A microscope was used to read the CWT codes then were recorded.


## **2.3 OVERNIGHT RETENTION CHECKS**

On Mondays and Tuesdays, up to 50 tagged fish at each site were placed in a holding bin to check for overnight CWT retention. This retention rate was then applied to the fish tagged during the entire week, so consistency in terms of tagger and methodology was maintained for the week. Any smolts that did not retain their tags were re-tagged. Overnight retention mortality was monitored, and any mortalities were investigated to determine possible causes for death and tag placement. All smolts held for overnight tag retention were then released in a calm location downstream.

## **2.4 DATA ANALYSIS**

Data was compiled daily and entered into Microsoft Excel spreadsheets. Analysis focused on summarizing counts by species, site, size, correctly estimating CWT releases by tag code and estimating the out-migrating smolt population using a mark-recapture. Environmental variables such as stream discharge and temperature were plotted to assess trends and to identify potential associations with migration activity.

A subsample of aged wild coho were plotted with a linear regression between fork length measurement and outmigration timing to assess the relationship between age, size, and run timing. However, most samples were taken in the beginning of the season resulting in an uneven representation of late samples. Therefore, an age-length key was not created. To assess the relationship between weight and length, a log-log linear regression model was applied for comparison to other years and streams. Fulton’s Condition Factor (K) was calculated to assess body condition, using the formula: K = W/L3 x 105, where W is weight in grams and L is fork length in millimeters. The CWT retention rates for the subsample of smolts held overnight at each site on Monday and Tuesday nights were averaged and applied to smolts tagged over the entire week. From the subsample, a small number of fish with unretained CWTs were retagged, and these fish were added back to the total tally of tagged smolts after retention rates were applied for each week.

An assumption of mark-recapture models is the equal probability of release and recapture, which is likely violated in an outmigrating smolt population where smolts move directionally from the system (Schwarz and Taylor 1998). Therefore a weekly stratified mark-recapture estimator was used to calculate the outmigrating smolt population (>70 mm) upon arrival to the upstream trap using SPAS (Schwarz 2025). A simple Lincoln-Petersen estimator (i.e. fully pooled strata) was compared to a fully stratified estimator (i.e. all weekly strata kept as-is), and an automatic logical pooling of strata estimator (i.e. strata pooled where catchability was similar) using AICc (Burnham and Anderson 2002).

Otherwise, all data manipulation and visualization were conducted using R Studio (v 4.4.2; R Core Team 2025) and packages tidyverse (Wickham et al., 2019), readr (Wickham et al., 2024), ggplot2 (Wickham, 2016), dplyr (Wickham et al., 2023), knitr (Xie, 2024) (Xie, 2015) (Xie, 2014), kableExtra (Zhu, 2024), scales (Wickham et al., 2023), tidyr (Wickham et al., 2024), flextable (Gohel and Skintzoas, 2024), FSA (Ogle et al., 2025), magrittr (Bache and Wickham, 2022), and nnet (Venables and Ripley, 2002)



\newpage

# **3. RESULTS AND DISCUSSION**

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 6.5,
  fig.height = 4.5,
  out.width = "120%"
)
```

```{r load-libraries, include=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(knitr)
library(tidyverse)
library(kableExtra)
library(scales)
library(tidyr)
library(flextable)
library(FSA)
library(magrittr)
library(dplyr)
library(nnet)
library(readxl)
```

```{r load-data, include=FALSE}
file_path <- "C:/Users/KIMS/Desktop/tbc report 2024/toboggan_smolt_dataentry_FINAL 2024.csv"
df_fish <- read_csv(file_path)

source("Smolt.R") # KP's scripts
source("Environmental.R")
```


## **3.1 ENVIRONMENTAL CONDITIONS **

The season began on May 1st, 2024 with only five nights of trapping per week, then switched to seven nights per week as of the trap set on the evening of May 19th, and back to five nights on June 14th. Trapping was only occasionally impacted by environmental conditions; during periods of very low or very high water, the traps experienced periods of partial or full inoperability due to debris or flow levels for two nights on June 18th and 19th (**Figure 2**). This period represents about 3.64% of the total trapping days/nights for the entire season. 

```{r Figure 2 flow, fig.cap="Figure 2. Total daily catch at both traps compared to daily discharge (in cubic meters per second (CMS) measured by the Hydrologic Station at Simpson Creek at the Mouth (08EE012; Environment Canada 2025). The Simpson Creek station was used as a surrogate for Toboggan Creek which did not have a station measuring discharge levels. The two creeks had similar patterns of water levels based on DFO staff observations and readings (data on file, DFO Smithers).", dpi=300}

plot.daily.co.smolt

```

Water temperatures increased throughout the season at a rate of 0.06 degrees Celsius per day (p<0.001), and were generally cooler in the mornings and warmer in the evenings (**Figure 3**). The Fence site appeared to be consistently cooler compared to the Red Barn site but the difference was not significant (p = 0.18).

```{r Figure 3 temps, fig.cap="Figure 3.Water temperature was measured at the start of each site visit for both the Red Barn (pink) and Fence (teal) sites. The “pull” points (circles) represent temperatures taken when traps were checked (usually in the morning when smolts were collected for processing). The “set” points (triangles) represent temperatures measured when traps were reset in the evenings.", dpi=300}

plot.water.temp24

```

## **3.2 TRAP CATCH**

A total of 18,615 coho smolts were caught during the 2024 season, and 20 of these were suspected to be hatchery releases from the previous spring based on their healed-over adipose scars, their lack of a secondary mark, and their sizes (**Table 1**). There were more coho (wild) caught at the Fence site than the Red  Barn site, with a total count of 12,864 and 5,731 smolts, respectively. The suspected hatchery smolts were sacrificed to retrieve their CWT codes and all but 3 of these fish were aged with scale samples. Of these fish, 12 had retained their CWTs, and 10 were successfully identified as 2023 Toboggan Creek hatchery releases from the 2021 brood year. The remainder were aged as 3 year olds and were assumed to be hatchery releases from the 2020 brood year based on their similar size, healed adipose scar, and lack of a secondary mark. The average fork length of these presumed hatchery coho was 104.30 mm (+/- 13.75 SD) and the average weight was 13.75 g (+/- 8.50 SD). Steelhead or rainbow trout (*O. mykiss*) were the most prevalent juvenile bycatch in the RSTs and only a small number of other species were observed (**Table 1**). Adult bycatch was released without counting.

```{r Table 1 fish count, fig.cap= "Table 1. Total count of fish for each juvenile species.", include=T}
df_fish <- df_fish %>%
  mutate(
    count = as.numeric(count),
    date = as.Date(date, tryFormats = c("%Y-%m-%d", "%m/%d/%Y")),
    mort = tolower(ifelse(is.na(mort), "no", mort)),
    sacrifice = tolower(ifelse(is.na(sacrifice), "no", sacrifice)),
    species = recode(species,
      "co-w" = "Coho (wild)",
      "st/rb" = "Steelhead or Rainbow trout",
      "co-a" = "Coho (hatchery)",
      "bt/dv" = "Bull trout or Dolly varden",
      "ct" = "Cutthroat",
      "mw" = "Mountain whitefish",
      "ch" = "Chinook",
      "pk" = "Pink"
    )
  )

summary_table <- df_fish %>%
  filter(!is.na(species) & species != "co-f") %>%
  group_by(species) %>%
  summarise(`Total Count` = sum(count, na.rm = TRUE), .groups = "drop") %>%
  rename(Species = species) %>%
  mutate(Species = factor(Species, 
                          levels = c("Coho (wild)", "Coho (hatchery)", 
                                     "Steelhead or Rainbow trout", 
                                     "Bull trout or Dolly varden", 
                                     "Cutthroat", 
                                     "Mountain whitefish", 
                                     "Chinook", 
                                     "Pink"))) %>%
  arrange(Species)

summary_table %>%
  kable(caption = "Table 1. Total count of fish for each juvenile species.", format = "pandoc",full_width=FALSE, latex_options = "HOLD_position",bootstrap_options = c("striped", "hover")) 
  # kable_styling(full_width = TRUE, position = "center", bootstrap_options = c("striped", "hover"))
```

**Note that these were hatchery fish released in previous years, identified by their healed-over adipose clip scar in the field and confirmed with dissected tag codes.*

## **3.3 SMOLT MIGRATION TIMING**

The peak for wild coho smolt outmigration in 2024 occurred between the last week of May and the first two weeks of June (**Figure 4**). The highest count of coho smolts was observed on June 4th with a total count of 1,681 smolts  across both sites. The lowest count of coho smolts was observed on June 24th,  with a total count of 17 smolts.

```{r Figure 4 timing, fig.cap="Figure 4. Daily count of coho (wild) at each site (Red Barn vs Fence). The upstream Red Barn hosted a 4 ft-diameter RST and the downstream  Fence site hosted a 6 ft-diameter RST.", dpi=300}
daily_counts_cow <- df_fish %>%
  filter(species == "Coho (wild)", site %in% c("barn", "fence")) %>%
  group_by(date, site) %>%
  summarise(daily_count = sum(count, na.rm = TRUE), .groups = "drop") %>%
  arrange(date)

if (nrow(daily_counts_cow) == 0) stop("Error: No data available for CO-W at Barn and Fence sites.")

# Plot daily counts for CO-W
ggplot(daily_counts_cow, aes(x = date, y = daily_count, color = site, linetype = site)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.5) +
  scale_x_date(
    breaks = seq(min(daily_counts_cow$date, na.rm = TRUE), max(daily_counts_cow$date, na.rm = TRUE), by = "1 week"),
    date_labels = "%d-%b"
  ) +
  labs(
    x = "Date",
    y = "Daily Count"
  ) +
  theme_bw() +
   theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## **3.4 MORTALITIES**

While great care and precaution was taken to ensure minimal harm to fish, a relatively small number of mortalities did occur, both accidentally and intentionally (**Table 2**). Accidental mortalities could have occurred through stress from the traps, anesthesia, handling, or a combination of these factors. Intentional mortalities occurred for two reasons: first, a small sample (0.17%) of wild coho were sacrificed to check for correct CWT placement to determine the correct head mold size to use for tagging. Second, all hatchery coho were culled in order to retrieve their CWT codes.

```{r Table 2 mortalities}
mortality_summary <- df_fish %>%
  filter(species != "co-f" & species != "Coho (hatchery)") %>% 
  group_by(species) %>%
  summarise(
    `Total Fish per Species` = sum(count, na.rm = TRUE),  
    `Total Accidental` = sum(ifelse(mort == "yes", count, 0), na.rm = TRUE),
    `Total Sacrifice` = sum(ifelse(sacrifice == "yes", count, 0), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    `% Accidental` = round((`Total Accidental` / `Total Fish per Species`) * 100, 2),
    `% Sacrifice` = round((`Total Sacrifice` / `Total Fish per Species`) * 100, 2),
    `% Accidental` = paste0(replace_na(`% Accidental`, 0), "%"),
    `% Sacrifice` = paste0(replace_na(`% Sacrifice`, 0), "%")
  ) %>%
  filter(`Total Accidental` > 0 | `Total Sacrifice` > 0)  # Remove species with zero mortality

# Add Coho (hatchery) data (all sacrificed)
coho_hatchery_data <- df_fish %>%
  filter(species == "Coho (hatchery)") %>%
  summarise(
    species = "Coho (hatchery)",
    `Total Fish per Species` = sum(count, na.rm = TRUE),
    `Total Accidental` = 0,
    `Total Sacrifice` = sum(count, na.rm = TRUE),
    `% Accidental` = "0%",
    `% Sacrifice` = "100%",
    .groups = "drop"
  )

# Bind Coho (hatchery) data and update column names and row order
mortality_summary <- bind_rows(mortality_summary, coho_hatchery_data) %>%
  rename(Species = species) %>%
  arrange(match(Species, c("Coho (wild)", "Coho (hatchery)")), Species)

# Display the table
mortality_summary %>%
  kable(format = "pandoc", booktabs = TRUE, align = c("l", "r", "r", "r", "r", "r"),
        col.names = c("Species", 
                      "Total Fish", 
                      "Total Accidental", 
                      "Total Sacrifice", 
                      "% Accidental", 
                      "% Sacrifice"),
        caption = "Table 2. Accidental and sacrificed mortality by species or wild/hatchery status. Percentages are calculated from the total of captured fish for each species."
  )
      
```

## **3.5 AGE AND SIZE**

**Figure 5** illustrates the log-log transformed weight-length relationship for measured wild coho. The linear trend describes the growth pattern in coho, where weight increases at an exponential rate relative to length. Fulton’s Condition Factor (K) was calculated and the mean condition factor was 1.08 (+/- 0.11 SD), with values ranging from 0.60 to 1.65.

```{r Figure 5, fig.cap="Figure 5. Log-log weight-length individual values and relationship for coho smolts. The scatterplot presents individual observations of fork length (mm) and weight (g), with a fitted regression line overlaid in blue and the regression described in the formula within the plot.", results = 'hide', dpi=300}
df_fish_coho <- df_fish %>% 
   filter(
    species %in% c("Coho (wild)"),  # Keep only co-w 
    !is.na(fork_length_mm), 
    !is.na(weight_g), 
    weight_g > 0, 
    fork_length_mm > 0
  )

length_weight_model <- lm(log10(weight_g) ~ log10(fork_length_mm), data = df_fish_coho)
summary(length_weight_model)

eq.log10FLwt <- paste0(expression(log[10](Weight)),"=",round(coef(length_weight_model)[1],2),
                       "+",round(coef(length_weight_model)[2],2),"*",expression(log[10](Length)))


ggplot(df_fish_coho, aes(x = fork_length_mm, y = weight_g)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", col = "blue") +
  scale_x_log10() +  # Apply log scale to x-axis
  scale_y_log10() +  # Apply log scale to y-axis for consistency
  # annotate("text",label= eq.log10FLwt,
  #          x = max(df_fish_coho$fork_length_mm) * 0.9,
  #          y = min(df_fish_coho$weight_g) * 3)
  labs(
    x = expression(log[10](Fork~Length~(mm))),
    y = expression(log[10](Weight~(g)))
  ) +
  annotate("text", 
           x = max(df_fish_coho$fork_length_mm) * 0.9,  # Leave in original scale
           y = min(df_fish_coho$weight_g) * 3,  # Leave in original scale
           label = eq.log10FLwt, 
           size = 3.5,  # Make text smaller
           hjust = 1, 
           fontface = "italic") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

#ggsave("co_w_size_histogram.png", width = 7, height = 5)
```

Based on association with age-verified smolts through scale sampling, a large proportion of the population sampled appeared to be two year olds (brood year 2022), followed by a proportion of three and even four year olds (brood years 2021 and 2020, respectively; **Figure 6**). Fish were sampled throughout the season and substantial growth occurred across every age. Sampling was unevenly concentrated towards the beginning of the season and most late samples were unrepresented, so no age-length keys were created. The majority of the largest smolts  were seen during the middle of the trapping season, while smaller smolts were seen more consistently across the season. 

```{r Figure 6, echo=FALSE, fig.cap="Figure 6. Subsample of aged wild coho plotted with fork length measurements throughout the season. Grey dots are unaged fish whose fork lengths were measured but no age sample was taken. Lines for each age are linear regressions with standard error shading. Only fish greater than 60 mm were measured. Age is represented by the Gilbert-Rich (GR) age designation system.", dpi=300}

plot.fork.length.age

```

## **3.6 CODED-WIRE TAGGING**

Wild coho smolts were sorted into size categories to be tagged using the correct head mold size. Coho between 86-115 mm were most abundant, with a total of 13,571 smolts intercepted (**Figure 7**). Coho between 70-85 mm and 116-190 mm were intercepted in similar numbers, with a total of 1,152 and 1,144 smolts, respectively. Coho between 60-69 mm and >190 mm were least abundant, with totals of 364 and 1 smolt intercepted, respectively.

```{r Figure 7, fig.cap="Figure 7. Total counts of coho (wild) by size category, which matched the head mold sizes used for tagging using CWTs.", dpi=300}
size_order <- c("little", "small", "big", "bigger", "biggest")
co_w_size_counts <- df_fish %>%
  filter(species == "Coho (wild)", !is.na(size_cwt)) %>%
  group_by(size_cwt) %>%
  summarise(total_count = sum(count, na.rm = TRUE), .groups = "drop") %>%
  mutate(size_cwt = factor(size_cwt, levels = size_order))

if (nrow(co_w_size_counts) == 0) stop("Error: No size data for Coho (wild) available.")

ggplot(co_w_size_counts, aes(size_cwt, total_count, fill = size_cwt)) +
  geom_bar(stat = "identity", color = "black") +
  labs(
    x = "Size Category", 
    y = "Total Coded-Wire Tags Applied",
    # fill = "Size Category"  # Change legend title # KP: legend just repeated x axis, removed
  ) +
  scale_x_discrete(
    labels = c(
      "little" = "60-69 mm",
      "small" = "70-85 mm",
      "big" = "86-115 mm",
      "bigger" = "116-190 mm",
      "biggest" = ">190 mm"
    )
  ) +
  scale_fill_manual(
    values = c(
      "little" = "lightblue", 
      "small" = "blue", 
      "big" = "darkblue", 
      "bigger" = "purple", 
      "biggest" = "red"
    ),
    labels = c(
      "little" = "60-69 mm",
      "small" = "70-85 mm",
      "big" = "86-115 mm",
      "bigger" = "116-190 mm",
      "biggest" = ">190 mm"
    )
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  scale_y_continuous(breaks = pretty_breaks(n = 6))
#ggsave("co_w_size_histogram.png", width = 7, height = 5)
```

24-hour CWT retention rates were generally quite high, and the average retention of the two representative days at the start of the week was applied to all injected fish that week (**Table 3**). Any non-retained tags from the subsample were re-tagged and re-added to the total releases. A total of 15,420 wild coho smolts were estimated to be released to the ocean with a CWT. Size classes were generally consistent within the coded-wire tag codes, so some anecdotal information about the size of fish in marine harvests can be drawn from the tag codes.
```{r Table 3 CWT tags}

kable(CWT.tag.summary2024 %>% select(-est.retained),
         col.names = c("Tag Code","#Inj","Est.Lost","#Re-tagged","Total Releases","Mean FL(mm)","SD FL(mm)","Mean Wt(g)","SD Wt(g)"),
         caption = "Table 3. Summary of the number of injected smolts for each coded-wire tag code series. Coho were generally sorted by size for coded-wire tagging. Estimated retention was based on a subsample of tagged smolts each week held for 24 hours after initial tagging. Average and standard deviation of fork length (FL in mm) and weight (Wt in g) for each tag code are also presented.",
         format = "pandoc", align = "c") 
  
```

## **3.7 MARK-RECAPTURE**

Weekly stratification for this project used two marks: upper and lower caudal fin clips, alternating each week on Monday. This stratification assumed that all fish that arrived at the upstream (Red barn) trap passed through the downstream (Fence) trap area within one week. Based on the first tags released from the Red Barn site each week, fish appeared to arrive at the Fence site by the next day, with recaptures  decreasing throughout the week (**Figure 8**). A small number of marks were not recognizable as either top or bottom caudal clips (n=6) and were removed from the mark-recapture analysis. While it appears that perhaps not all fish had moved past the downstream trap in one week, our observations across both trapping sites support that the majority of them did.

```{r Figure 8 timing, fig.cap="Figure 8. Comparison of daily mark sequence applied at the Red Barn (the upstream trap) in the bubble plot below the x-axis scaled by number of applied tags. The bar chart shows the daily number of recaptures at the Fence site (the downstream trap) with upper and lower caudal clips. “None” refers to a caudal clip which cannot be distinguished as either top or bottom, which was the case for a small number of recaptures (n=6) throughout the season. These fish were removed from the mark-recapture analysis."}

 plot.mark.timing24bubble

```

Marks by week and their recaptures during the following two weeks are summarized in **Table 4** following Schwarz (2025).

```{r Table 4 MR strata}
kable(x,
      caption = "Table 4. Summary of weekly releases and recoveries by strata following Schwarz (2025). Row and column names are isoweeks, with rows (s) representing the release week from the Red Barn site and columns (t) representing the recapture week at the Fence site. The final row (s+1) represents the total number of unmarked smolts captured at the fence that week, and the final column (t+1) is the total number of marked smolts from the barn that week that were never recaptured."
)
```

A comparison of fully pooled Lincoln-Petersen, fully stratified, and logically-pooled mark-recapture estimators indicated that a logically pooled estimator was the most parsimonious model with the lowest AICc (**Table 5**). The first two weeks (19, 20) and the last two weeks (25 and 26) were pooled based on their similar catch probabilities and smaller number of recaptures (full model output in available in **Appendix A**). Using this estimator, the population of wild coho smolts (>70 mm) migrating from Toboggan Creek in 2024 was  `r paste0(report.table[1,"Nhat"]," +/- ",report.table[1,"Nhat.se"],"SE")`.

```{r Table 5 MR}

kable(report.table,
      col.names = c("Model","s","t","AICc","Delta AICc","gof.chisq", "gof.df", "gof.p",  "Nhat", "Nhat SE"),
        caption = "Table 5. Comparison of candidate stratified population estimate models for the Toboggan Creek Smolt project. The Full weekly strata model did not pool any strata, the Auto (logical) pooling model pooled the first two and the last two weeks for the tagging season, and the Logical pooling to single row (LP) pooled all rows to essentially reduce to a Lincoln-Petersen estimate. s and t are the number of weekly strata of tagging and recapture, respectively, AICc is Akaike’s Information Criterion corrected for small sample sizes, GOF is a goodness of fit test, in this case a Chi2, df is degrees of freedom and p is probability. N ̂ is the population estimate of wild smolts and N ̂ SE is the standard error around the estimate.
", 
        format = "pandoc",full_width=T, latex_options = "HOLD_position",bootstrap_options = c("striped", "hover")
)

```

\newpage

# **4. CONCLUSION**

Following the pause in hatchery smolt releases due to a disease outbreak in 2024, the Smithers North Coast Stock Assessment Division was presented with a unique opportunity to focus on a wild smolt mark-recapture. The 2024 Toboggan Smolt Coho Smolt Program generated a wild coho smolt population estimate of 35,717 with a CV under 5%. Though age sampling was unevenly distributed through the season, we were able to characterize the size distribution and age range within the coho smolt population, including the identification of a substantial number of 3 and 4 year old smolts that appeared to have reared in the watershed for far longer than the expected 1-2 year rearing period. A total count of 18,595 wild coho smolts were caught and tagged. Of that, tag retention observations showed that 15,420 smolts were estimated to be released to the ocean with a CWT. While Toboggan Creek Hatchery released an annual average of about 32,000 tagged smolts from 1987 to 2023 (data on file, DFO Smithers), the potentially higher survivability of wild smolts may compensate for their lower release numbers (Unpublished data, DFO Smithers, 2024). With some adjustments to future years of sampling, the program procedures piloted this year met and exceeded its objectives.

The continuation of the wild smolt marking program will be key for evaluating the assumption that wild and hatchery coho survival is comparable, which has been a key assumption of the Toboggan Creek Coho Indicator Program. This project provided valuable information  about the wild coho smolt population in Toboggan Creek, which can be compiled with the many years of smolt data for Toboggan Creek to assess trends and changes over time. Information about marine survival should be available after the fall of 2025, when the majority of the smolts released in spring of 2024 should return as spawning adults (PSC NBTC, 2022).



\newpage

# **5.	ACKNOWLEDGEMENTS**

Funding for this project was provided by the Pacific Salmon Commission’s Northern Endowment Fund and the Department of Fisheries and Oceans Canada (DFO). Field work was conducted by staff from the DFO and the Toboggan Creek Salmon and Steelhead Enhancement Society (TCSSES). The pilot project operating procedures were developed by V. Chicatun and K. Peck, with feedback from R. Whitmore. Thank you to reviewers…. XXX

Thanks to everyone who contributed to the project this year, and especially to the landowners who generously provided access to the creek and trap infrastructure through their properties.

\newpage

# **6.	REFERENCES**

Bache, S.M. and Wickham, H. (2022) magrittr: A forward-pipe operator for R. R package version 2.0.3. Available at: https://CRAN.R-project.org/package=magrittr

Burnham, K.P. and Anderson, D.R. (2002) Model selection and multimodel inference: A practical information-theoretic approach. 2nd edn. New York, NY: Springer.

Chicatun, V. and Peck, K. (2024) Toboggan smolt standard operating procedures 2024 – Draft V2. Smithers, BC: Department of Fisheries and Oceans Canada, North Coast Stock Assessment Branch.

Environment Canada (2025) Real-time hydrometric data graph for Simpson Creek at the mouth (08EE012) [BC]: Water level and flow. Government of Canada. Available at: https://wateroffice.ec.gc.ca/

Fisheries and Oceans Canada, North Coast Stock Assessment Branch (2002) Northern British Columbia coho salmon information summary. Pacific Salmon Commission Technical Report No. 47.

Gohel, D. and Skintzos, P. (2024) flextable: Functions for tabular reporting. R package version 0.9.7. Available at: https://CRAN.R-project.org/package=flextable

MacLellan, S.E. (2004) Guide for sampling structures used in age determination of Pacific salmon. Fisheries and Oceans Canada. Available at: https://waves-vagues.dfo-mpo.gc.ca/library-bibliotheque/342153.pdf

Noble, C.A.J., Kingshott, S., Morven, N., McCulloch, C., Davies, S., Alexander, R.F. and Bocking, R.C. (2020) Adult and juvenile coho salmon (Oncorhynchus kisutch) enumeration and coded-wire tag recovery analysis for Ksi Ts’oohl Ts’ap (Zolzap Creek), British Columbia, 2011–2019. Fisheries and Oceans Canada, Stock Assessment Division.

Ogle, D.H., Doll, J.C., Wheeler, A.P. and Dinno, A. (2025) FSA: Simple fisheries stock assessment methods. R package version 0.9.6. Available at: https://CRAN.R-project.org/package=FSA

Pacific Salmon Commission, Northern Boundary Technical Committee (PSC NBTC) (2002) Status of coho salmon stocks and fisheries in the Northern Boundary Area. Report TCNB (02)-3.

R Core Team (2025) R: A language and environment for statistical computing. Vienna, Austria: R Foundation for Statistical Computing. Available at: https://www.R-project.org/

Schwarz, C.J. (2025) SPAS: Stratified-Petersen analysis system. R package version 2025.2.1. Available at: https://CRAN.R-project.org/package=SPAS. https://doi.org/10.32614/CRAN.package.SPAS

Schwarz, C.J. and Taylor, C.G. (1998) Use of the stratified-Petersen estimator in fisheries management: estimating the number of marked fish in the presence of unequal capture probabilities. Canadian Journal of Fisheries and Aquatic Sciences. https://doi.org/10.1139/f98-104

Shaul, L.D., Crabtree, K.F., Kemp, M. and Olmsted, N. (2009) Coho salmon studies at Hugh Smith Lake, 1982–2007. Fishery Manuscript Series No. 09-04. Alaska Department of Fish and Game.

SKR Consultants (2001) Toboggan Creek coho smolt enumeration 2001. Available at: https://www.gov.bc.ca/TobogganCreekCohoSmoltEnumeration2001_1519664118144_9663846798.pdf

Van Will, P., Stiff, H.W., Miyagi, E. and Duncan, K. (2010) Enumeration of juvenile and adult coho salmon at Black Creek, Vancouver Island, 2006. Canadian Manuscript Report of Fisheries and Aquatic Sciences No. 2928. Fisheries and Oceans Canada, Pacific Region.

Venables, W.N. and Ripley, B.D. (2002) Modern applied statistics with S. 4th edn. New York, NY: Springer.

Wickham, H. (2016) ggplot2: Elegant graphics for data analysis. New York, NY: Springer.

Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T.L., Miller, E., Bache, S.M., Müller, K., Ooms, J., Robinson, D., Seidel, D.P., Spinu, V., Takahashi, K., Vaughan, D., Wilke, C., Woo, K. and Yutani, H. (2019) Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686. https://doi.org/10.21105/joss.01686

Wickham, H., François, R., Henry, L., Müller, K. and Vaughan, D. (2023) dplyr: A grammar of data manipulation. R package version 1.1.4. Available at: https://CRAN.R-project.org/package=dplyr

Wickham, H., Hester, J. and Bryan, J. (2024) readr: Read rectangular text data. R package version 2.1.5. Available at: https://CRAN.R-project.org/package=readr

Wickham, H., Pedersen, T.L. and Seidel, D. (2023) scales: Scale functions for visualization. R package version 1.3.0. Available at: https://CRAN.R-project.org/package=scales

Wickham, H., Vaughan, D. and Girlich, M. (2024) tidyr: Tidy messy data. R package version 1.3.1. Available at: https://CRAN.R-project.org/package=tidyr

Wolf, P. (1951) A trap for the capture of fish and other organisms moving downstream. Transactions of the American Fisheries Society, 80, pp. 41–45.

Xie, Y. (2014) knitr: A comprehensive tool for reproducible research in R. In: Stodden, V., Leisch, F. and Peng, R.D. (eds.) Implementing reproducible computational research. Boca Raton, FL: Chapman and Hall/CRC.

Xie, Y. (2015) Dynamic documents with R and knitr. 2nd edn. Boca Raton, FL: Chapman and Hall/CRC.

Xie, Y. (2024) knitr: A general-purpose package for dynamic report generation in R. R package version 1.49. Available at: https://yihui.org/knitr/

Zhu, H. (2024) kableExtra: Construct complex table with “kable” and pipe syntax. R package version 1.4.0. Available at: https://CRAN.R-project.org/package=kableExtra

\newpage

# **Appendix A**

Figure A1: The most parsimonious model output from SPAS (Schwarz 2025) following logical pooling of strata.
```{r SPAS model results}

SPAS::SPAS.print.model(mod3)

```


```{r Figure X-bubble plot, results = 'hide', fig.cap="Figure X. Bubble plot representation of the observed age-length key for wild coho smolts at Toboggan Creek. Age is represented by the Gilbert-rich (GR) age designation system.", echo=F, include=F}
# 
# df_co_w <- df_fish %>% 
#   filter(species == "Coho (wild)") %>% 
#   select(fork_length_mm, `GR AGE`)
# 
# df_co_w <- df_co_w %>% mutate(`GR AGE` = ifelse(`GR AGE` == "", NA, `GR AGE`))
# df_co_w <- df_co_w %>% mutate(lcat5 = lencat(fork_length_mm, w = 5))
# 
# aged <- df_co_w %>% filter(!is.na(`GR AGE`))
# unaged <- df_co_w %>% filter(is.na(`GR AGE`))
# 
# aged <- aged%>% mutate(`GR AGE` = as.numeric(`GR AGE`))
# 
# aged <- aged %>% filter(`GR AGE` %in% c(22, 33, 44))
# 
# alk.freq <- xtabs(~ lcat5 + `GR AGE`, data = aged)
# rowSums(alk.freq)
# alk <- prop.table(alk.freq, margin=1)
# 
# ggplot(aged, aes(x = fork_length_mm, y = `GR AGE`)) +
#   geom_point(
#     color = "skyblue",
#     size = 3,
#     alpha = 0.7
#   ) +
#   scale_y_continuous(
#     breaks = c(22, 33, 44),
#     labels = c("22", "33", "44"),
#     expand = expansion(mult = c(0.05, 0.05))
#   ) +
#   scale_x_continuous(
#     breaks = seq(50, 200, 25),
#     expand = expansion(mult = c(0.02, 0.02))
#   ) +
#   labs(
#     x = "Fork Length (mm)",
#     y = "Age"
#   ) +
#   theme_minimal(base_size = 12) +
#   theme(
#     panel.grid.major = element_line(color = "grey80", linewidth = 0.4),
#     panel.grid.minor = element_blank(),
#     axis.text = element_text(color = "black"),
#     axis.title = element_text(color = "black"),
#     axis.ticks = element_line(color = "black"),
#     axis.ticks.length = unit(0.2, "cm")
#   )

```

